---
title: "AlexLinearModelling - STDS"
author: "Alex Brooks"
date: "9/25/2018"
output: html_document
---
###Modelling merged dataset
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(reshape2)
library(caret)
library(glmnet)
library(mlbench)
library(ROCR)
```

##Linear modelling using the Immunization with Everything inc tax

```{r}

#import data from Ethan's final merge which includes immunisation coverage, postcode, PHN, Electorate, SEIFA deciles and ranks and mean taxable income per postcode
everything <- read.csv("../cleaned_data/immunization_with_everything_taxation_update.csv")
str(everything)

View(everything)

```

```{r}
# * mergeddata - the source data frame we want to clean

# Returns - a data frame

mergeddata <- everything %>%
    select('state.x', 'postcode', 'year', 'age', 'pc_immun_class', 'PHN_number','Usual.resident.population','Rank.within.Australia...Decile', 'Electoral.division', 'mean_tax')

#selected these variables to mix postcode, age, immunisation classes, PHN, usual population, SEIFA rank within Australia in deciles, SEIFA score, Electorate and party details and mean tax of postcode.


#turn categorical variables into factors
mergeddata$postcode <- as.factor(mergeddata$postcode)
mergeddata$year <- as.factor(mergeddata$year)
mergeddata$age <- as.factor(mergeddata$age)
mergeddata$pc_immun_class <- as.factor(mergeddata$pc_immun_class)
str(mergeddata)
```

```{r}
#cleaning the data
#Filter on age
#remove pc_immun(duplicates pc_immun_class)
#remove score (duplicates Rank.withn.Australia ... Decile)
#remove PartyNms but keep electorate
#discard caution for now

#NPs are useless to us so let's remove them, this means all class 0 won't be in this model

mergeddata <- mergeddata %>%
  filter(pc_immun_class != 0)
```
##Let's look at the age groups
```{r}
one_year <- filter(mergeddata, age==1)
two_year <- filter(mergeddata, age==2)
five_year <- filter(mergeddata, age==5)

#THe ordinal package expects the dependent variable to be an ordered factor
one_year$pc_immun_class <- ordered(one_year$pc_immun_class)

#Divide mean_tax by a thousand to get lower numbers (polr doesn't seem to like the big numbers)
one_year <- transform(one_year, mean_tax_thousands = mean_tax/1000)

str(one_year)

set.seed(42)
 
#split test and train 60:40
inTrain <- createDataPartition(y = one_year$pc_immun_class, p = .60, list = FALSE)
training <- one_year[inTrain,]
testing <- one_year[-inTrain,]
#use 3000 rows of data as training to stop computer timing out while we build a model
experiment <- head(training, 3000)

dim(experiment)
dim(testing)
```
## Commentary around goals of prediction
I'm trying to predict PC_immun_class in 2017-18 or 2018-19 (with the last year of data in our set from 2016-17)
The variables that could help build the predictive model could be:
- Mean tax (changes each year)
- Age group (changes each year)
- Electoral division 
- SEIFA rank by decile
- SEIFA score
- Postcode
- Usual resident population
- PHN number
- State

It's likely this requires ordinal regression, as our target variable has an order. pc_immun_class is an ordinal predictor with 8 different ordered classes. 0 = NP, which we will remove in the first instance. We may have to impute them later to balance the model.
```{r}
#see pc_immun_class counts by filtered age groups.
#for one year olds
ggplot(one_year)+geom_bar(aes(x=pc_immun_class))
# 6 and 7 are the dominant class for one year olds, with 8 close in count to 5
```
```{r}
#for two year olds
ggplot(two_year)+geom_bar(aes(x=pc_immun_class))
# there are higher counts in class 5, but 6 & 7 are still the dominant class, with less in 8
```
##Age could be a predictor for future years
```{r}
#for five year olds
ggplot(five_year)+geom_bar(aes(x=pc_immun_class))
#5 year old has more even distribution across 6, 7 and 8.
```

```{r}
#Ordinal regression model - because pc_immun_class has multiple categories and they have a specific order

library(MASS)

model1 <- polr(pc_immun_class ~ Rank.within.Australia...Decile + mean_tax_thousands, data = training, Hess = TRUE)

#let's try the ordinal library
#library(ordinal)
#model <- clm(pc_immun_class ~ Rank.within.Australia...Decile + PHN_number, data = training)

summary(model1)

```
Now let's try to assess model 1
```{r}
coefs <- coef(model1)

coefs
```

```{r}
# Find the p-value for model 1's t-value of 13.545
pt(13.545, 400-3, lower.tail=FALSE)*2
```
##Try a second model, adding PHN_number
```{r}
model2 <- polr(pc_immun_class ~ Rank.within.Australia...Decile + mean_tax_thousands + PHN_number, data = training, Hess = TRUE)

#let's try the ordinal library
#library(ordinal)
#model <- clm(pc_immun_class ~ Rank.within.Australia...Decile + PHN_number, data = training)

summary(model2)
```

```{r}
coefs <- coef(model2)

coefs
```
##Try a third model, adding the continuous variable of usual resident population  
```{r}
model3 <- polr(pc_immun_class ~ Rank.within.Australia...Decile + mean_tax_thousands + PHN_number + Usual.resident.population, data = training, Hess = TRUE)

#let's try the ordinal library
#library(ordinal)
#model <- clm(pc_immun_class ~ Rank.within.Australia...Decile + PHN_number, data = training)

#summary(model3)
#when I introduce large numbers like "usual.resident.population, I get "Error in svd(x) : infinite or missing values in 'x'

```
## QUESTIONS:
- do we need to impute the missing NAs?
- what steps do we need to take next to get this model to predict future immun_classes?
- how do I assess the fit against the test set?
